Generalized Tractability

### 2023-04-21 GeneralizedTractability_230418.tex
Updates from Overleaf

Certain common notions of tractability are covered by the general situation described above.
{\small 
\begin{equation*}
	\begin{array}{r@{\qquad}c@{\qquad}cc}
		\text{Tractability type} & T(\varepsilon^{-1},d,\vp)
		& T(0,d,\vp)\\
		\toprule
		\text{strong polynomial algebraic} & \max(1,\varepsilon^{-p}) & 1\\
		\text{polynomial algebraic} & \max(1,\varepsilon^{-p})d^{q} & d^{q}\\
		\text{strong polynomial exponential} &  [\max(1,\log(1 + \varepsilon^{-1})]^p & 1\\
		\text{polynomial exponential } &
		[\max(1,\log(1 + \varepsilon^{-1}))]^p  d^{q} & d^{q} \\
        \text{quasi-polynomial} & 
        \exp\{p(1+\log(\max(1,\varepsilon^{-p}))(1+\log(d))\} &
        \exp\{p(1+\log(d))\}
	\end{array}
\end{equation*}}
Here above, $\vp = (p,q)$ in some cases and $\vp$ is a scalar in other cases. A general reference 
for results on algebraic tractability is \cite{NW08}, for exponential tractability we refer to \cite{HKW19} and \cite{KW19}.

@@

Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,p)= (\max\{\varepsilon^{-1},1\})^p 
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, p \ge 0.
\]
Then Theorem \ref{thm_main_strong_tract2} yields the  equivalent condition for strong algebraic polynomial tractability: 
\[
 \sup_{d\in\naturals} \sum_{i=L_p}^\infty (\min\{\lambda_{i,d},1\})^p < \infty \text{ for some } p \ge 0, L_p \in \naturals.
\]
Furthermore, the optimal $p$ is the infimum of all $p$ for which the above condition holds.
This essentially recovers the result in \cite[Theorem 5.1]{NW08}.
\end{example}

@@

\begin{theorem}\label{thm:subhT}
	Let $T$ be a tractability function as specified in \eqref{eq:Tspec} and satisfying \eqref{eq:Tconditions}.  Moreover, let $\lim_{d \to \infty} T(\varepsilon^{-1},d,\vp) = \infty$ for all $\varepsilon > 0$, $\vp > \vzero$.  The problem is sub-$h$-$T$ tractable if and only if
	\[
	S_c = S_{c,\vp} :=\sup_{d\in\naturals}  \, \sum_{i=1}^\infty \frac1{h(c\, T(\lambda_{i,d}^{-1},d,\vp))} <\infty \qquad \forall c > 0.
	\]
\end{theorem}
\begin{proof}
We suppress the $\vp$ dependence, because it does not enter into the proof.

@@

Since $S_c \ge 1/h(cT(0,1))$ by definition, it follows that $S_c \, h(cT(0,1)) \, h(cT(\varepsilon^{-1},d))  \ge 1$ and so 
\begin{align*}
\max(1,\comp(\varepsilon,d)) 
& \le  \lceil  S_c \, h(cT(0,1)) \, h(cT(\varepsilon^{-1},d))   \rceil 
 =  \lceil 
(\widetilde{S}_c/2) \, h(cT(\varepsilon^{-1},d))  
\rceil \\
& \le   \widetilde{S}_c \, h(cT(\varepsilon^{-1},d))  
\qquad\forall c, \varepsilon > 0, \, d \in \naturals.
\end{align*}
Since $h$ is strictly increasing, so is $h^{-1}$, and thus
\begin{align*}
h^{-1}(\max(1,\comp(\varepsilon,d)) ) & \le  h^{-1} \left(  
\widetilde{S}_c \, h(cT(\varepsilon^{-1},d))  \right) 
\\
&  \le  h^{-1} (2 S_c) + cT (\varepsilon^{-1},d).
\quad \forall c, \varepsilon > 0, \, d \in \naturals \qquad \text{by \eqref{eq:submult_h_inverse}.}
\end{align*}

@@

Note that 
\[
\varepsilon^{-1}_{\max}(d,c) = \max(V_c - d,\lambda_{\max}^{-1}) 
= \begin{cases} V_c - d \le V_c, & d \in \{1, \ldots, d^*_c : = \lfloor V_c - \lambda_{\max}^{-1} \rfloor\} \\
\lambda_{\max}^{-1}, & d \in \{d^*_c+1, \ldots\}
\end{cases}
\]
Recalling that $T$ is non-decreasing in its arguments, this implies that 
\begin{align*}
    \MoveEqLeft{\sup_{d\in\naturals} \,
		 \frac{ h(cT(\varepsilon^{-1}_{\max}(d,c),d))}{h(2cT(\lambda_{\max}^{-1},d))}}\\
  &\le \max \left \{  
  \sup_{d \in \{1, \ldots, d^*_c\}} \,
		 \frac{ h(cT(\varepsilon^{-1}_{\max}(d,c),d))}{h(2cT(\lambda_{\max}^{-1},d))}, 
  \sup_{d\in \{d^*_c+1, \ldots \}} \,
		 \frac{ h(cT(\varepsilon^{-1}_{\max}(d,c),d))}{h(2cT(\lambda_{\max}^{-1},d))}
  \right \} \\
  & \le \max \left \{  
  \sup_{d \in \{1, \ldots, d^*_c\}}  \,
		 \frac{ h(cT(V_c,d))}{h(2cT(\lambda_{\max}^{-1},d))}
  , 
  \sup_{d\in \{d^*_c+1, \ldots \}}  \,
		 \frac{ h(cT(\lambda_{\max}^{-1},d))}{h(2cT(\lambda_{\max}^{-1},d))}
  \right \}
  \\
  & \le \max \left \{  
		 \frac{ h(cT(V_c,d^*_c))}{h(2cT(\lambda_{\max}^{-1},1))}, 
  \sup_{d\in \{d^*_c+1, \ldots \}}  \,
		 \frac{ h(cT(\lambda_{\max}^{-1},d))}{[h(cT(\lambda_{\max}^{-1},d))]^2}
  \right \} \\
  & = \max \left \{  
		 \frac{ h(cT(V_c,d^*_c))}{h(2cT(\lambda_{\max}^{-1},1))}, 
		 \frac{1}{h(cT(\lambda_{\max}^{-1},d_c^*+1))}
  \right \} \\
  & < \infty.
\end{align*}
Combining this bound with \eqref{eq:lastone} establishes the finiteness of $S_{2c}$ and completes the proof of the necessary condition.

This completes the proof.
\end{proof}

\begin{example}[Weak Tractability]
	Suppose that 
	\[
	T(\varepsilon^{-1},d,\vp) = \varepsilon^{-s} + d^t, \quad \forall \varepsilon > 0, \ d \in \naturals, \quad \text{where } \vp = (s,t), \qquad h(x) = \exp(x) \quad \forall x \ge 0.
	\]
	Then Theorem \ref{thm:subhT} implies that 
	\[\lim_{\varepsilon^{-1}+d\rightarrow \infty} \frac{\log( \max(1,\comp(\varepsilon,d)))}{\varepsilon^{-s}+d^t} = 0 
	\iff 
	 \sup_{d\in\naturals}  \, \sum_{i=1}^\infty \exp(-c\, (\lambda_{i,d}^{-s} + d^t )) <\infty \qquad \forall c > 0.
	\]
	The expression on the left is the notion of $(s,t)$-weak tractability \cite{??}, and the condition on the right was derived in \cite{??}.  Weak tractability corresponds to $(s,t)=(1,1)$, and uniform weak tractability corresponds to the case when a problem is weakly tractable for all positive $s$ and $t$.
\end{example}

>>>>

Certain common notions of tractability are covered by the general situation described above.
{\small 
\begin{equation*}
	\begin{array}{r@{\qquad}c@{\qquad}cc}
		\text{Tractability type} & T(\varepsilon^{-1},d,\vp)
		& T(0,d,\vp)\\
		\toprule
		\text{strong polynomial algebraic} & \max(1,\varepsilon^{-p}) & 1\\
		\text{polynomial algebraic} & \max(1,\varepsilon^{-p})d^{q} & d^{q}\\
		\text{strong polynomial exponential} &  [\max(1,\log(1 + \varepsilon^{-1})]^p & 1\\
		\text{polynomial exponential } &
		[\max(1,\log(1 + \varepsilon^{-1}))]^p  d^{q} & d^{q} \\
        \text{quasi-polynomial} & 
        \exp\{p(1+\log(\max(1,\varepsilon^{-1}))(1+\log(d))\} &
        \exp\{p(1+\log(d))\}
	\end{array}
\end{equation*}}
Here above, $\vp = (p,q)$ in some cases and $\vp$ is a scalar in other cases. A general reference 
for results on algebraic tractability is \cite{NW08}, for exponential tractability we refer to \cite{HKW19} and \cite{KW19}.

@@

Let the tractability function, $T$,
be defined by
\[
 T(\varepsilon^{-1},d,p)= (\max\{\varepsilon^{-1},1\})^p 
 \qquad \forall \varepsilon > 0, \,  d \in \naturals, \, p \ge 0.
\]
\fred{May we pull the exponent $p$ inside the $\max$?} Then Theorem \ref{thm_main_strong_tract2} yields the  equivalent condition for strong algebraic polynomial tractability: 
\[
 \sup_{d\in\naturals} \sum_{i=L_p}^\infty (\min\{\lambda_{i,d},1\})^p < \infty \text{ for some } p \ge 0, L_p \in \naturals.
\]
Furthermore, the optimal $p$ is the infimum of all $p$ for which the above condition holds.
This essentially recovers the result in \cite[Theorem 5.1]{NW08}.
\end{example}

@@

\begin{theorem}\label{thm:subhT}
	Let $T$ be a tractability function as specified in \eqref{eq:Tspec} and satisfying \eqref{eq:Tconditions}.  Moreover, let $\lim_{d \to \infty} T(\varepsilon^{-1},d,\vp) = \infty$ for all $\varepsilon > 0$, $\vp > \vzero$.  The problem is sub-$h$-$T$ tractable if and only if
	\[
	S_c = S_{c,\vp} :=\sup_{d\in\naturals}  \, \sum_{i=1}^\infty \frac1{h(c\, T(\lambda_{i,d}^{-1},d,\vp))} <\infty \qquad \forall c > 0.
	\]
\end{theorem}
\begin{proof}
The idea of this proof follows \cite{WW17}.
We suppress the $\vp$ dependence in the proof, because it insignificant.

@@

Since $S_c \ge 1/h(cT(0,1))$ by definition, it follows that $S_c \, h(cT(0,1)) \, h(cT(\varepsilon^{-1},d))  \ge 1$ and so 
\begin{align*}
\max(1,\comp(\varepsilon,d)) 
& \le  \lceil  S_c \, h(cT(0,1)) \, h(cT(\varepsilon^{-1},d))   \rceil 
 =  \lceil 
(\widetilde{S}_c/2) \, h(cT(\varepsilon^{-1},d))  
\rceil \\
& \le   \widetilde{S}_c \, h(cT(\varepsilon^{-1},d))  
\qquad\forall c, \varepsilon > 0, \, d \in \naturals.
\end{align*}
Since $h$ is strictly increasing, so is $h^{-1}$, and thus by \eqref{eq:submult_h_inverse}, 
\begin{align*}
h^{-1}(\max(1,\comp(\varepsilon,d)) ) & \le  h^{-1} \left(  
\widetilde{S}_c \, h(cT(\varepsilon^{-1},d))  \right) 
\\
&  \le  h^{-1} (2 S_c) + cT (\varepsilon^{-1},d).
\quad \forall c, \varepsilon > 0, \, d \in \naturals.
\end{align*}

@@

Note that 
\[
\varepsilon^{-1}_{\max}(d,c) = \max(V_c - d,\lambda_{\max}^{-1}) 
= \begin{cases} V_c - d \le V_c, & d \in \{1, \ldots, d^*_c : = \lfloor V_c - \lambda_{\max}^{-1} \rfloor\} \\
\lambda_{\max}^{-1}, & d \in \{d^*_c+1, \ldots\}
\end{cases}
\]
Recalling that $T$ is non-decreasing in its arguments, this implies that 
\begin{align*}
    \MoveEqLeft{\sup_{d\in\naturals} \,
		 \frac{ h(cT(\varepsilon^{-1}_{\max}(d,c),d))}{h(2cT(\lambda_{\max}^{-1},d))}}\\
  &\le \max \left \{  
  \sup_{d \in \{1, \ldots, d^*_c\}} \,
		 \frac{ h(cT(\varepsilon^{-1}_{\max}(d,c),d))}{h(2cT(\lambda_{\max}^{-1},d))}, 
  \sup_{d\in \{d^*_c+1, \ldots \}} \,
		 \frac{ h(cT(\varepsilon^{-1}_{\max}(d,c),d))}{h(2cT(\lambda_{\max}^{-1},d))}
  \right \} \\
  & \le \max \left \{  
  \sup_{d \in \{1, \ldots, d^*_c\}}  \,
		 \frac{ h(cT(V_c,d))}{h(2cT(\lambda_{\max}^{-1},d))}
  , 
  \sup_{d\in \{d^*_c+1, \ldots \}}  \,
		 \frac{ h(cT(\lambda_{\max}^{-1},d))}{h(2cT(\lambda_{\max}^{-1},d))}
  \right \}
  \\
  & \le \max \left \{  
		 \frac{ h(cT(V_c,d^*_c))}{h(2cT(\lambda_{\max}^{-1},1))}, 
  \sup_{d\in \{d^*_c+1, \ldots \}}  \,
		 \frac{ h(cT(\lambda_{\max}^{-1},d))}{[h(cT(\lambda_{\max}^{-1},d))]^2}
  \right \} \\
  & = \max \left \{  
		 \frac{ h(cT(V_c,d^*_c))}{h(2cT(\lambda_{\max}^{-1},1))}, 
		 \frac{1}{h(cT(\lambda_{\max}^{-1},d_c^*+1))}
  \right \} \\
  & < \infty.
\end{align*}
Combining this bound with the upper bound on $S_c$ in \eqref{eq:lastone} establishes the finiteness of $S_{2c}$ and completes the proof of the necessary condition.
This concludes the proof of Theorem \ref{thm:subhT}.
\end{proof}

\begin{example}[Weak Tractability]
	Suppose that 
	\[
	T(\varepsilon^{-1},d,s,t) = \max(1,\varepsilon^{-1})^s + d^t, \quad \forall \varepsilon,s,t > 0, \, d \in \naturals, \qquad h(x) = \exp(x) \quad \forall x \ge 0.
	\]
	Then Theorem \ref{thm:subhT} implies that 
	\[\lim_{\varepsilon^{-1}+d\rightarrow \infty} \frac{\log( \comp(\varepsilon,d))}{\varepsilon^{-s}+d^t} = 0 
	\iff 
	 \sup_{d\in\naturals}  \, \sum_{i=1}^\infty \exp(-c\, (\min(1,\lambda_{i,d})^{s} + d^t )) <\infty \qquad \forall c > 0.
	\]
	The expression on the left is the notion of $(s,t)$-weak tractability \cite{??}, and the condition on the right was derived in \cite{??}.  Weak tractability corresponds to $(s,t)=(1,1)$, and uniform weak tractability corresponds to the case when a problem is weakly tractable for all positive $s$ and $t$.
\end{example}

\section{Summary}

\fred{Do we consider the case of normalized error?}
